---
title: "MATH 392 Power Calculations"
author: "EJ Arce"
date: "9 March 2018"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 6
---

```{r, include=FALSE}
require(tidyverse) # load packages
require(resampledata)
set.seed(30) # set seed for reproducibility
```

Given $\mu = 100, \sigma^2 = 15^2, \bar{X} = 105, \alpha = .05$, for what sizes n will power be .7? .75? .9?

$$
H_{0}: \mu = 100
$$

$$
H_{A}: \mu > 100
$$

## Power = .7

Under each hypothesis, the sampling distributions will follow

$$
X_{H_0} \sim N(\mu,\sigma/\sqrt{n})
$$

$$
X_{H_A} \sim N(\bar{X}_{obs},\sigma/\sqrt{n})
$$

With $\alpha=.05$, the critical value $x_{crit}$at which we reject the null hypothesis is $\mu$ +1.645($\sigma/\sqrt{n}$). $\beta$ is the area under the alternative hypothesis distribution to the left of this critical value. With $\beta$ = .3, we can refer to the standard normal distribution to find how many standard errors away from the mean our critical value is:

```{r}
qnorm(.3,0,1)
```

Thus,

$$
x_{crit} = 100 + 1.645\frac{\sigma}{\sqrt{n}} = 105 - .5244\frac{\sigma}{\sqrt{n}}
$$

Now solve for n:

$$
\begin{aligned}
(1.645+.5244) \frac{\sigma}{\sqrt{n}} & = 5 \\
(\frac{(2.189)15}{5})^2 & = n \\
n & = 42.35
\end{aligned}
$$

Thus for a power of at least .7, the sample size must be at least 43.

## Power = .75

Generalizing the equations for $x_{crit}$,

$$
100 + 1.645\frac{\sigma}{\sqrt{n}} = 105 - q_{z_{\beta}}\frac{\sigma}{\sqrt{n}}
$$

where $q_{z_{\beta}}$ is the quantile for the standard normal distribution under which the area to the left of the quantile is $\beta$. Thus for $\beta$ = .25, $q_{z_{\beta}}$ is

```{r}
qnorm(.25,0,1)
```

Thus

$$
\begin{aligned}
(1.645+.6745) \frac{\sigma}{\sqrt{n}} & = 5 \\
(\frac{(2.319)15}{5})^2 & = n \\
48.4 & = n
\end{aligned}
$$

A sample size of at least 49 is needed for a power of .75

## Power = .9

```{r}
qnorm(.1,0,1)
```

$$
\begin{aligned}
(1.645+1.282) \frac{\sigma}{\sqrt{n}} & = 5 \\
(\frac{(2.927)15}{5})^2 & = n \\
77.11 & = n
\end{aligned}
$$

A sample size of at least 78 is needed for a power of .9.

