---
title: "MATH 392 Problem Set 2"
author: "EJ Arce"
date: "2 February 2018"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 6
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE)
require(tidyverse) # load packages
require(resampledata)
options(digits=4) # print 4 decimal places only
set.seed(30) # set seed for reproducibility
```

## 3.4

### a
We are testing to see if the distributions of the proportion of flights delayed by more than 20 minutes, $\theta$, differs by airline:
$$
H_0: \theta_{UA} - \theta_{AA} = 0
$$
  
$$
H_A: \theta_{UA} - \theta_{AA} \neq 0
$$

```{r}
FlightDelays <- FlightDelays %>%
  mutate(Delayed20 = ifelse(Delay > 20, 1, 0))

# Calculating observed difference in the two groups' mean proportions
xobs <- mean(FlightDelays$Delayed20[FlightDelays$Carrier == "UA"] -
               mean(FlightDelays$Delayed20[FlightDelays$Carrier=="AA"]))

# Running simulation for hypothesis testing
nsim <- 10000
means <- rep(NA, nsim)
for(i in 1:nsim){
  perm <- sample(FlightDelays$Carrier, replace=F)
  means[i] <- mean(FlightDelays$Delayed20[perm=="UA"]) - 
                     mean(FlightDelays$Delayed20[perm=="AA"])
}
simdf <- data.frame(means)

# Plotting the simulated null distribution
ggplot(simdf, aes(x = means)) +
  geom_density() +
  geom_vline(xintercept = xobs, col = "red") +
  ggtitle("Mean differences from 10000 simulations")
```

The red vertical line indicates the difference in proportions observed in the actual dataset. The p-value for a two-tailed test is calculated below.

```{r}
(sum(means > xobs) + 1)/(length(means) + 1) * 2
```

### b

We are testing to see if the variance in flight delays for United Airlines is greater than the variance for American Airlines. Thus, we are conducting a one-tailed significance test. Speicically,


$$
H_0: \rho_{UA} \leq \rho_{AA}
$$
  
$$
H_A: \rho_{UA} >\rho_{AA}
$$

```{r}
# Variance in flight delay lengths for each carrier
varUA <- var(FlightDelays$Delay[FlightDelays$Carrier == "UA"])
varAA <- var(FlightDelays$Delay[FlightDelays$Carrier == "AA"])
varUA
varAA
```

varUA and varAA indicate the variances of United Airlines' and American Airlines' delay times, respectively. A simulation just like the last problem will be used to test if the difference in these variances is statistically significant.

```{r}
# Calculate test statistic
obs.diff <- varUA-varAA

# Run simulation
nsim <- 10000
diffs <- rep(NA, nsim)
for(i in 1:nsim){
  perm <- sample(FlightDelays$Carrier, replace=F)
  diffs[i] <- var(FlightDelays$Delay[perm=="UA"]) - 
                     var(FlightDelays$Delay[perm=="AA"])
}
simdf <- data.frame(diffs)

# Plotting the simulated null distribution
ggplot(simdf, aes(x = diffs)) +
  geom_density() +
  geom_vline(xintercept = obs.diff, col = "red") +
  ggtitle("Variance differences from 10000 simulations")
```

```{r}
(sum(diffs > obs.diff) + 1)/(length(diffs) + 1)
```

The density plot and calculated p-value show that the observed variance for United Airlines delays is not significantly greater than observed variances for American Airlines.

## 3.16

### a

```{r}
table(GSS2002$Gender,GSS2002$Pres00)
```

### b

```{r}
gender <- GSS2002$Gender
pres <- GSS2002$Pres00
chisq.test(gender,pres)
```

### c

```{r}
# Remove NAs
GSS2002 <- GSS2002 %>%
  filter(!is.na(Gender),
         !is.na(Pres00))
# Chi-squared test using permutations
x2.obs <- chisq.test(gender,pres)$statistic
nsim <- 10000
x2.stats <- rep(NA,nsim)
for(i in 1:nsim){
  perm <- xtabs(~sample(Gender, replace=F) + Pres00, data = GSS2002)
  x2.stats[i] <- chisq.test(perm)$statistic
}
(sum(x2.stats > x2.obs) + 1)/(nsim+1)
```

None of the simulated $\chi^{2}$ values were greater than our observed $\chi^{2}_{obs}$ value of 33.29, resulting in our very low p-value.

## 3.22

```{r}
q <- c(.2,.4,.6,.8)
d <- data.frame("quantile" = q)
obs.stats <- c(12.57,16.87,20.73,24.66)
d<-cbind(d,obs.stats)
exp.stats <- rep(NA,4)
for(i in 1:4){
  exp.stats[i] <- qnorm(q[i],22,7)
}
d <- cbind(d,exp.stats)
d
# Calculate observed chi-squared value
((3.54^2)/16.11) + ((3.36^2)/20.23) + ((3.04^2)/23.77) + ((3.23^2)/27.89)
```

## 3.31

### Empirical Solution

```{r}
nsim <- 10000
t.obs <- rnorm(nsim,0,1)
ts <- data.frame(t.obs)
ts <- ts %>%
  arrange(t.obs) %>%
  mutate(t.obs = abs(t.obs))
p.val <- rep(NA,nsim)
for(i in 1:nsim){
  p.val[i] <- (sum(ts$t.obs>ts$t.obs[i])+1)/(length(ts$t.obs)+1)
}
ts <- cbind(ts, p.val)
ggplot(ts, aes(x=p.val)) +
  geom_density()
```

As we'd expect, the simulated density plot follows a uniform distribution.

### Analytical Solution
Consider a test statistic t = $T(x_{1},...,x_{n})$. Its corresponding p-value is calculated by solving
$$
p = Pr(T(X) \leq t|H_{0}).
$$
This makes p a random variable as well, since its calculated probability depends on the random variable T(X). Thus the p-value p follows some probability distribution $p = F_{T}(T)$. Since the p-values are drawn from the distribution of T(X), then the p-values have a one-to-one correspondence to each observed test statistic t. Thus,
$$
F_{P}(p) = Pr(P \leq p) = Pr(T(X) \leq t) = Pr(F_{T}(T) \leq p).
$$

$$
F_{P}(p) = Pr(F^{-1}_{T}F_{T}(T) \leq F^{-1}_{T}(p))
$$

$$
F_{P}(p) = Pr(T \leq F_{T}^{-1}(p))
$$

$$
F_{P}(p) = F_{T}(F^{-1}_{T}(p)) = p
$$

This shows that P(T) follows a uniform distribution, where Pr(p) = 1/n.

## 3.32

Let Z denote the standard normal random variable. Then Z ~ N(0,1). Suppose $X = Z^{2}$. Show that X ~ $\chi^{2}_{df=1}$.

The pdf of Z is already known to be
$$
f_{Z}(z) = \frac{1}{\sqrt{2\pi}}e^{-\frac{z^2}{2}}
$$

Using the cdf method,

$$
F_{X}(x) =
P(X \leq x) =
P(Z^{2} \leq x) =
P(-\sqrt{x} \leq Z \leq \sqrt{x}) =
F_{Z}(\sqrt{x}) - F_{Z}(-\sqrt{x})
$$
$$
P(-\sqrt{x} \leq Z \leq \sqrt{x}) =
F_{Z}(\sqrt{x}) - F_{Z}(-\sqrt{x}) =
\int_{-\sqrt{x}}^{\sqrt{x}}
\frac{1}{\sqrt{2\pi}}
e^{\frac{-x^2}{2}}
dx
$$

$$
F_{Z}(\sqrt{x}) - F_{Z}(-\sqrt{x}) =
2\int_{0}^{\sqrt{x}}
\frac{1}{\sqrt{2\pi}}
e^{\frac{-x^2}{2}}
dx
$$

$$
F_{Z}(\sqrt{x}) - F_{Z}(-\sqrt{x}) = 2 \frac{1}{2}erf (\frac{\sqrt{x}}{\sqrt{2}})
$$

$$
F_{Z}(\sqrt{x}) - F_{Z}(-\sqrt{x}) = erf (\frac{\sqrt{x}}{\sqrt{2}})
$$

$$
f_{X}(x) =
\frac{\partial}{\partial x} (F_{Z}(\sqrt{x}) - F_{Z}(-\sqrt{x}))=
2f_{Z}(\sqrt{x}) \frac{1}{2}x^{-\frac{1}{2}}
$$

$$
f_{X}(x) =
\frac{1}{\sqrt{2\pi}}e^{-\frac{x}{2}} x^{-\frac{1}{2}}
$$

$$
f_{X}(x) = \frac{x^{-\frac{1}{2}}e^{-\frac{x}{2}}}{\sqrt{2\pi}}
$$

Notice that 
$$\Gamma(1/2) =
\int_{0}^\infty t^{-1/2}e^{-t} dt=
\sqrt{\pi},$$
so we get

$$
f_{X}(x) = \frac{x^{-\frac{1}{2}}e^{-\frac{x}{2}}}{\sqrt{2} \Gamma(1/2)}
$$

Thus,

$f_{X}(x)$ ~ $\chi^2_{df = 1}$