---
title: "MATH 392 Problem Set 3"
author: "EJ Arce"
date: "7 February 2018"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 6
---


```{r, include=FALSE}
require(tidyverse) # load packages
require(resampledata)
set.seed(30) # set seed for reproducibility
```

## 4.8

$$
n = 20, \mu = 6, \sigma^{2} = 10
$$

$$
P(\bar{X} \leq 4.6) =
P(\frac{\bar{X}-\mu}{\sigma / \sqrt{n}} \leq
\frac{4.6 -\mu}{\sigma / \sqrt{n}})
$$

$$
P(\bar{X} \leq 4.6) =
P(\frac{\bar{X}-6}{\sqrt{10} / \sqrt{20}} \leq
\frac{4.6 -6}{\sqrt{10}  / \sqrt{20}})
$$

```{r}
# Calculate Z score of interested statistic
z.obs <- (4.6-6)/(sqrt(10)/sqrt(20))
z.obs
```

$$
P(\bar{X} \leq 4.6) = P(Z \leq -1.98)
$$

```{r}
# Calculate probability using the cdf of N(0,1)
pnorm(z.obs, 0, 1)
```

$$
P(\bar{X} \leq 4.6) = .02385
$$

## 4.9

$$
f_{X}(x) = \frac{3}{16}(x-4)^{2} for 2 \leq 6
$$

Find E[X]:

$$
E[X] = \int_{2}^{6} x \frac{3}{16}(x-4)^{2} dx
$$

$$
E[X] = \int_{2}^{6} \frac{3}{16}x(x^{2}-8x+16) dx
$$

$$
E[X] = \int_{2}^{6} \frac{3}{16}x^{3}- \frac{3}{2}x^{2}+3x dx
$$

$$
E[X] = \frac{3}{64}x^{4} - \frac{1}{2}x^{3} + \frac{3}{2}x^{2} \vert_{2}^{6}
$$

$$
E[X] = 4
$$

Find V[X]:

$$
V[X] = E[X^{2}] - E[X]^{2}
$$

We already calculated that $E[X]=4$, so $E[X]^{2}$ = 16. Now solve for $E[X^2]$:

$$
E[X^{2}] = \int_{2}^{6} x^{2} f(x) dx
$$

$$
E[X^{2}] = \int_{2}^{6} x^{2} \frac{3}{16}(x-4)^{2} dx
$$

$$
E[X^{2}] = \int_{2}^{6} \frac{3}{16}x^{4}- \frac{3}{2}x^{3}+3x^{2} dx
$$

$$
E[X^{2}] = \frac{3}{80}x^{5}- \frac{3}{8}x^{4}+x^{3} \vert_{2}^{6}
$$

```{r}
# Calculate
e.xsq <- (3*(6^5)/80 - 3*(6^4)/8 +6^3) -
  (3*(2^5)/80 - 3*(2^4)/8 +(2^3))
e.xsq
sq.ex <- 4^2
var.x <- e.xsq-sq.ex
sd.x <- sqrt(var.x)
sd.x
```

Thus $V[X] = 2.4$, so SD[X] = $\sqrt{2.4}$ = 1.549. Now, $n = 244, \mu = 4, \sigma^{2} = 2.4$.

$$
P(\bar{X} \geq 4.2) =
P(\frac{\bar{X}-\mu}{\sigma / \sqrt{n}} \geq
\frac{4.2 -\mu}{\sigma / \sqrt{n}})
$$

$$
P(\bar{X} \geq 4.2) =
P(\frac{\bar{X}-4}{\sqrt{2.4} / \sqrt{244}} \geq
\frac{4.2 -4}{\sqrt{2.4} / \sqrt{244}})
$$

$$
P(\bar{X} \geq 4.2) =
P(Z \geq \frac{4.2 -4}{\sqrt{2.4} / \sqrt{244}})
$$

```{r}
# Calculate Z score of interested statistic
z.obs <- (4.2-4)/(sqrt(2.4)/sqrt(244))
z.obs
```

```{r}
# Calculate probability using cdf from N(0,1)
1 - pnorm(z.obs,0,1)
```


## 4.12

### a

Let X be a random sample of size 30 from the exponential distribution with rate $\lambda = .1$. The expected value of the sample mean is the same as the expected value of the population, by linearity of expectation. Thus, E[X] = $\frac{1}{\lambda}$ = 10

### b

```{r}
# Run simulation
nsim <- 1000
n <- 30
rate <- 1/10
means <- rep(NA,nsim)
for(i in 1:nsim){
  sample <- rexp(n,rate)
  means[i] <- mean(sample)
}
sum(means >= 12)/nsim
```

### c

Since 12.1% of the samples had means of 12 or greater, this observation is not that unusual.

## 4.13

### a

Since X ~ $N(20,8^{2})$ and Y $~ N(16,7^{2})$ are independent variables, and W = $\bar{X} + \bar{Y}$, then W ~ $N(36,\frac{8^{2}}{10} + \frac{7^{2}}{15})$

### b

```{r}
nsim <- 10000
w <- rep(NA,nsim)
for(i in 1:nsim){
  x <- rnorm(10,20,8)
  y <- rnorm(15,16,7)
  w[i] <- mean(x) + mean(y)
}
# Compute mean and standard error
mean(w)
sd(w)
# Plot sampling distribution
w <- data.frame(w)
ggplot(w,aes(x=w)) +
  geom_density() +
  geom_vline(xintercept=40)
```

### c

```{r}
sum(w<40)/nsim
```


## 4.18



## 4.20



## 4.21



