---
title: "MATH 392 Problem Set 5 (Corrected)"
author: "EJ Arce"
date: "9 March 2018"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 6
---

```{r, include=FALSE}
require(tidyverse) # load packages
require(resampledata)
set.seed(30) # set seed for reproducibility
```

## 6.39

{$\hat{\beta}_n$} is consistent if and only if $\lim_{n\rightarrow\infty} P(|\hat{\beta}_n - \beta| < \varepsilon) = 1$ $\forall$ $\varepsilon$ > 0.

We know that for any n, the distribution of $X_{max}$ follows the distribution

$$
F_{X_{max}}(x) = \begin{cases}
1,x > \beta \\
(\frac{x}{\beta})^n, 0<x<\beta \\
0, x < 0 \\
\end{cases}
$$

Thus

$$
\begin{aligned}
\lim_{n\rightarrow\infty} P(|\hat{\beta}_n - \beta| < \varepsilon)
& = \lim_{n\rightarrow\infty} P(\beta - \varepsilon < \hat{\beta}_n < \beta + \varepsilon) \\ 
& = \lim_{n\rightarrow\infty} P(\beta - \varepsilon < X_{max} < \beta)  \\
& = \lim_{n\rightarrow\infty} (\frac{x}{\beta})^n |_{\beta-\varepsilon}^{\beta}\\
& = 1 - \lim_{n\rightarrow\infty} (\frac{\beta-\varepsilon}{\beta})^n \\
& = 1 \forall \varepsilon >0
\end{aligned}
$$

Thus {$\hat{\beta}_n$} is consistent.

## 6.40

{$\hat{\sigma}_n^2$} is consistent iff

$$
\lim_{n\rightarrow\infty} P(|\hat{\sigma}_n^2 - \sigma^2| < \varepsilon)
= 1 \Leftrightarrow 
\begin{cases}
\lim_{n\rightarrow\infty} bias[\hat{\sigma}_n^2] = 0 \\
\lim_{n\rightarrow\infty} V[\hat{\sigma}_n^2] = 0
\end{cases}
$$

First show that $\lim_{n\rightarrow\infty} bias[\hat{\sigma}_n^2] = 0$:

From Problem Set 4, 

$$
E[\hat{\sigma}_n^2] - \sigma^2 =
\frac{n-1}{n}\sigma^2 - \sigma^2 = 
\frac{-1}{n}\sigma^2
$$

Thus $\lim_{n\rightarrow\infty} E[\hat{\sigma}_n^2] - \sigma^2 = 0 \Rightarrow bias[\hat{\sigma}_n^2] \rightarrow 0$.

Now show that $\lim_{n\rightarrow\infty} V[\hat{\sigma}_n^2] = 0$:

From B.16, $\frac{\hat{\sigma}_n^2n}{\sigma^2} \sim \chi^2_{n}$, so $V[\frac{\hat{\sigma}_n^2n}{\sigma^2}] = 2n$ $\forall$ $n \epsilon \mathbb{N}$. Thus


$$
\begin{aligned}
\lim_{n\rightarrow\infty} (\frac{n}{\sigma^2})^2 V[\hat{\sigma}_n^2]
& = \lim_{n\rightarrow\infty} V[\frac{\hat{\sigma}_n^2n}{\sigma^2}] = \lim_{n\rightarrow\infty} 2n \\
\lim_{n\rightarrow\infty} V[\hat{\sigma}_n^2]& = \lim_{n\rightarrow\infty}\frac{2n}{(\frac{n}{\sigma^2})^2} \\ 
& = \lim_{n\rightarrow\infty} \frac{2\sigma^4}{n} \\
& = 0
\end{aligned}
$$

Thus {$\hat{\sigma}_n^2$} is consistent.

## 7.3

### a

$$
X \sim N(185, 50^2) , n = 100, \bar{X} = 210 
$$

$$
P(q_{.05} < \frac{\bar{X}-\mu}{\sigma/\sqrt{n}} < q_{.95}) = 1-\alpha = .9
$$

Solve for $\mu$:

$$
P( \, \bar{X} - q_{.95}\frac{\sigma}{\sqrt{n}} < 
\mu < 
\bar{X} - q_{.05}\frac{\sigma}{\sqrt{n}}) \, =
.9
$$

$$
(210 - 1.645\frac{50}{\sqrt{100}} < 
\mu < 
210 + .1.645\frac{50}{\sqrt{100}}) =
(201.8 < \mu < 218.8)
$$

Thus the 90% confidence interval is (201.8, 218.8).

### b

Solve for n:

$$
\bar{X} -LB_{.05} =
\bar{X} -(\bar{X}-1.96(\frac{50}{\sqrt{n}}))
\leq 10
$$

$$
1.96\frac{50}{10} \leq \sqrt{n}
$$

$$
n \geq 96.04
$$

A sample size of at least 97 is needed.

### c

Like b,

$$
2.576\frac{50}{10} \leq \sqrt{n}
$$

$$
n \geq 165.9
$$

A sample size of at least 166 is needed.

## 7.8

For n = 30, the code generating the simulation is below, along with the proportion of times the confidence interval missed $\mu$.

```{r}
# Run simulation
low <- 0
high <- 0
n <- 30
q <- qt(.975, n-1)
N <- 10^5
for(i in 1:N){
  x <- rgamma(n,shape = 5, rate = 2)
  xbar <- mean(x)
  s <- sd(x)
  L <- xbar - q*s/sqrt(n)
  U <- xbar + q*s/sqrt(n)
  if(U<5/2){
    low <- low + 1
  }
  if(5/2<L){
    high <- high + 1
  }
}

(high+low)/N
```

```{r, echo=F}
n <- 60
low <- 0
high <- 0
q <- qt(.975, n-1)
N <- 10^5
for(i in 1:N){
  x <- rgamma(n,shape = 5, rate = 2)
  xbar <- mean(x)
  s <- sd(x)
  L <- xbar - q*s/sqrt(n)
  U <- xbar + q*s/sqrt(n)
  if(U<5/2){
    low <- low + 1
  }
  if(5/2<L){
    high <- high + 1
  }
}
(high+low)/N
```

```{r, echo=F}
n <- 100
low <- 0
high <- 0
q <- qt(.975, n-1)
N <- 10^5
for(i in 1:N){
  x <- rgamma(n,shape = 5, rate = 2)
  xbar <- mean(x)
  s <- sd(x)
  L <- xbar - q*s/sqrt(n)
  U <- xbar + q*s/sqrt(n)
  if(U<5/2){
    low <- low + 1
  }
  if(5/2<L){
    high <- high + 1
  }
}
(high+low)/N
```

```{r, echo=F}
n <- 250
low <- 0
high <- 0
q <- qt(.975, n-1)
N <- 10^5
for(i in 1:N){
  x <- rgamma(n,shape = 5, rate = 2)
  xbar <- mean(x)
  s <- sd(x)
  L <- xbar - q*s/sqrt(n)
  U <- xbar + q*s/sqrt(n)
  if(U<5/2){
    low <- low + 1
  }
  if(5/2<L){
    high <- high + 1
  }
}

(high+low)/N
```

Compared to the next three numbers for n=60, n=100, and n=250, the frequency of missing $\mu$ converges to the thoretical frequency of .05. As n increases, the number of estimates that are too high increases, while the number of estimates that are too low decreases.

## 7.34

We are given

$$
X \sim Gamma(2,\lambda)
$$

$$
2\lambda x \sim \chi^2_{df=4},
f_{2\lambda X}(2\lambda x)
$$

Thus, the quantiles will be calculated in terms of $2\lambda x$, using the $\chi^2_{df=4}$  distribution.

$$
\begin{aligned}
.95 = P(q_{1,x} < \lambda < q_{2,x})
& =  P(q_{1,2\lambda x} < 2 \lambda X < q_{2,2\lambda x}) \\ 
& = P(\frac{q_{1,2\lambda x}}{2X} < \lambda < \frac{q_{2,2\lambda x}}{2X}) \\
& = P(\frac{.2422}{X} < \lambda < \frac{5.572}{X}) \\
\end{aligned}
$$

Thus the 95% confidence interval for $\lambda$ is $(\frac{.2422}{X},\frac{5.572}{X})$.