---
title: "MATH 392 Problem Set 7"
author: "EJ Arce"
date: "26 March 2018"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 6
---


```{r, include=FALSE}
require(tidyverse) # load packages
require(resampledata)
set.seed(30) # set seed for reproducibility
```

## 4

$$
Var(X) = 5
$$

$$
Var(Y) = 7
$$

$$
Cov(X,Y) = 2
$$

$$
\begin{aligned}
Var(2X-5Y) & = Var(2X) + Var(5Y) - 2(2)(5)Cov(X,Y) \\
& = 4Var(X) + 25Var(Y) - 20Cov(X,Y) \\
& = 20 + 175 - 40 \\
& = 155
\end{aligned}
$$

## 7

```{r}
corrExerciseB <- corrExerciseB
```

### a

```{r}
cov <- cov(corrExerciseB$X, corrExerciseB$Y)
sigx <- sd(corrExerciseB$X)
sigy <- sd(corrExerciseB$Y)
rho <- cov/(sigx*sigy)
rho
```

### b

```{r}
d <- aggregate(corrExerciseB[, 1:2], list(corrExerciseB$Z), mean)
d
```

### c

```{r}
ggplot(d, aes(x = X, y = Y)) +
  geom_point(aes(color = Group.1))
```

```{r}
cor(d$X,d$Y)
```

The correlation coefficient is much higher between means of X and Y than the correlation coefficient between each observation.

## 9

Show that $\sum_{i=1}^{n} y_{i} - \hat{y}_{i} = 0$.

$\sum_{i=1}^{n} y_{i} - \hat{y}_{i} = \sum_{i=1}^{n} y_{i} - (a + bx_{i})$, where a and b are found using the function $g(a,b) = \sum_{i=1}^{n} (y_{i} -(a+bx_{i}))^2$, setting $\frac{\partial g}{\partial a}$ = 0 and solving for a and b. Thus,

$$
g(a,b) = \sum_{i=1}^{n} y_{i}^2 -2y_{i}(a+bx_{i}) +a^2 +2abx_{i} + (bx_{i})^2
$$

$$
\begin{aligned}
\frac{\partial g}{\partial a} = 0 & = \sum_{i=1}^{n} -2y_{i} + 2a + 2bx_{i}\\
0 & = 2 \sum_{i=1}^{n} a + bx_{i} -y_{i} \\
0 & = 2na + 2 \sum_{i=1}^{n}bx_{i}-y_{i} \\
a & = \frac{1}{n}\sum_{i=1}^{n}y_{i}-bx_{i}
\end{aligned}
$$

Now plug a into the original equation:

$$
\begin{aligned}
\sum_{i=1}^{n} y_{i} - (a + bx_{i})
& = -na + \sum_{i=1}^{n} y_{i} - bx_{i} \\
& = -n\frac{1}{n}\sum_{i=1}^{n}y_{i}- bx_{i} + \sum_{i=1}^{n}y_{i}- bx_{i} \\
\sum_{i=1}^{n} y_{i} - \hat{y}_{i}& = 0
\end{aligned}
$$

## 14

```{r}
df <- Illiteracy
```

### a

```{r}
ggplot(df,aes(x=Illit, y = Births)) + geom_point()
```

At first glance, the scatterplot appears to show a positive relationship between female birth rate and illiteracy.

### b

```{r}
m1 <- lm(Births ~ Illit, data = df)
m1$coefficients
summary(m1)$r.squared
```

The least-squares equation is $\hat{y} = 1.9488 + .0545x$. A 1% increase in the illteracy of women is associated with an increase of .0545 in the number of births per woman. With an $r^2$ of .5908, about 59% of the variance in the number of births per woman can be explaned by the variance in illteracy rate.

### c

```{r}
yhat <- predict(m1)
resid.d <- data.frame(residuals = m1$residuals,
                      y = yhat)
ggplot(resid.d, aes(x = yhat, y = residuals)) + geom_point() + geom_hline(yintercept = 0)
```

The residuals are randomly scattered when plotted against the predicted values of births. There is no clear linearity nor outliers, so this straight-line model is appropriate.

### d

Although the scatterplot, regression model, and residual plot all imply a positive relationship between births and illiteracy rate, we cannot conclude that the variance in one varaible causes the variance in another variable. For example, it is possible that a third variable causes the variance in both births and illiteracy.
