---
title: "MATH 392 Problem Set 6"
author: "EJ Arce"
date: "19 March 2018"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 6
---


```{r, include=FALSE}
require(tidyverse) # load packages
require(resampledata)
set.seed(30) # set seed for reproducibility
```

```{r}
# Load data
x <- c(47, 126, 285, 318, 142, 55, 231, 
       102, 164, 85, 242, 62, 289, 290)
```

## 1.1

$$
X \sim Unif(0,\theta)
$$

$$
H_{0}: \theta = 345
$$

$$
H_{A}: \theta > 345
$$

$$
\hat{\theta}_{MLE, corr} = \frac{n + 1}{n} X_{max}
$$

```{r}
# Calculate test statistic
xmax <- max(x)
n <- length(x)
theta.test.mle <- (n+1)*xmax/n
```

```{r}
# Simulate sampling distribution of thetas from null distribution
nsim <- 10^5
max.null <- 345
theta.mle <- rep(NA,nsim)
for(i in 1:nsim){
  sample <- runif(n,0,max.null)
  theta.mle[i] <- (n+1)*max(sample)/n
}
null.samp.mle <- data.frame(theta.mle)
ggplot(null.samp.mle, aes(x=theta.mle)) +
  geom_density() +
  geom_vline(xintercept = theta.test.mle, col = "red")
```

Now, calculate the proportion of simulations that produced estimates more extreme than the test statistic:

```{r}
(sum(null.samp.mle$theta.mle > theta.test.mle) + 1)/(length(null.samp.mle$theta.mle) + 1)
```

Using a reasonable $\alpha$ level, the simulation results are not statistically significant enough to reject the null hypothesis.

## 1.2

```{r}
theta.test.mom <- 2*mean(x)
nsim <- 10^5
max.null <- 345
theta.mom <- rep(NA,nsim)
for(i in 1:nsim){
  sample <- runif(n,0,max.null)
  theta.mom[i] <- 2*mean(sample)
}
null.samp.mom <- data.frame(theta.mom)
ggplot(null.samp.mom, aes(x=theta.mom)) +
  geom_density() +
  geom_vline(xintercept = theta.test.mom, col = "red")
```

```{r}
(sum(null.samp.mom$theta.mom > theta.test.mom) + 1)/(length(null.samp.mom$theta.mom) + 1)
```

## 1.3

The two types of error committed when concluding hypothesis tests are

* Type I: rejecting $H_{0}$ when it is actually true
* Type II:retaining $H_{0}$ when $H_{A}$ is actually true

In the case of the German tank problem, committing a type I error would mean that the Western allies believe that there are more German tanks than there actually are. A type II error would mean that the Western allies believe the Germans have 345 tanks, when in fact they have more. Thus, a type II error is more consequential than a type I error.

## 1.4

Suppose $\alpha$ = .05.

```{r}
n <- 17
nsim <- 10^5
theta1 <- 325
theta2 <- 335
theta3 <- 345
theta4 <- 355
theta5 <- 365
```

```{r}
# Power curve for theta parameters
f.mle <- function(theta.a){
  nsim <- 10^5
  theta.0 <- 345
  test <- (n+1)*(max(x))/n
  thetas.0 <- rep(NA,nsim)
  thetas.a <- rep(NA,nsim)
  for(i in 1:nsim){
    sample.0 <- runif(n,0,theta.0)
    sample.a <- runif(n,0,theta.a)
    thetas.0[i] <- (n+1)*(max(sample.0))/n
    thetas.a[i] <- (n+1)*(max(sample.a))/n
  }
  sampling.0 <- data.frame(thetas.0)
  sampling.a <- data.frame(thetas.a)
  alpha.crit <- quantile(sampling.0$thetas.0,.95)
  power <- sum(sampling.a$thetas.a > alpha.crit)/length(sampling.a$thetas.a)
  power
}
f.mle(theta5)
```

```{r}
d <- data.frame(x = c(theta1,theta2,theta3,theta4,theta5),
                y = c(f.mle(theta1),f.mle(theta2),
                      f.mle(theta3),f.mle(theta4),f.mle(theta5)))
ggplot(d,aes(x=x,y=y)) + geom_point()
```

As expected, as the true alternative hypothesis increases, the power increases.
